### 2019-06-25 thesis-template/msc-thesis.tex
update

\chapter*{Abstract}
%\input{abstract}
Many problems in science and engineering require
tensor field representations. Tensor fields, are most commonly associated with stress distributions (cf. Cauchy-stress tensor) in 2D/3D grids, but also have some other meanings in practice in physics. By means of global illumination techniques, we motivate a new method to visualize tensor fields, which is also capable of visualizing the light transport (propagation) in a topologically defined scene geometry. As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles and is able to approximate light distributions for given light source position(s) and direction(s) until convergence. The transmission profiles within this grid are considered as crystal fiber structures as an underlying physical model (, eg., gemstones: tiger's eye's cat's eye effect). The consequent task is to model anisotropic fibre structures with the orientation and anisotropy measures of the underlying tensor field in the grid.
For this, we derive an eigenbasis by PCA (principal component analysis) for every cell of the tensor field and form an ellipsoid equation as transmission (transfer/weighting) function for the transmitted light profiles. Hence, the resulting light distribution can be visualized via 2D scalar field color coding or polar plots for directional information. Additionally, we measure impulse responses of the tensor field with Dirac-pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions. This map is considered as a global illumination energy flow field.
We gain inspiration by vector field particle tracing's FTLE approach for analyzing the gradient of the flow field of the resulting light distributions. Similar to how the FTLE detects ridges/LCS (Lagrangian Coherent Structures) in vector fields (flow fields), our approach captures these structures likewise, namely attracting/repelling/saddling tensor field lines (TFL) which frequently represent key structures in nature, in tensor fields. We denote this entity a light transport gradient (LTG) which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global light transport  directed by imprint/engraving in crystelline fiber structures modeled by the tensor field's ellipsoid glyhps. We evaluate our approaches for plausibility and comparability within an extensive test campaign and demonstrate its performance on several examples of datasets.
\newpage

@@

The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies, e.g., clustering, structuring or projection techniques. When it comes to tensor fields, this comes in handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor-medical resonance imaging) \cite{hlawatsch}, when measuring water molecule diffusion characteristics within tissue or in simulated (or measured) stress tensor fields from solid and fluid mechanics. In these domains of science the problem arises, that tensor fields can be generated (simulated) or measured with physical/mathematical background, but not be regarded intuitively. Necessity begets ingenuity as always, which lets visualization methods and systems emerge in this context. Concurrently, the problem setting of this work encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids consisting of mainly second-order tensors (cf. Cauchy stress tensor). These grids span a spatial domain in general with sample positions indicated by vertices. Tensor fields feature a sampling grid consistent of tensors which is spaced uniformly in the commonly used Cartesian grid. A Cartesian grid, which we will utilize in this context, is a normalized uniform grid as a special case of a regular grid. We aim to use global illumination methods to model light transport directed by tensor fields within such type of grids, which we consider as light propagation volumes (LPVs). Congruently, the most basic objective of this work is to design a simple light propagation scheme yet following the most basic but crucial physical principles and laws, like energy conservation and propagation attenuation, for tensor field visualization. Furthermore, it should hold the additional possibility to model light transport in a geometrical scene. This propagation scheme should either be used to visualize the light transport paths (,defined by the anisotropy of tensor fields and tensor magnitudes of a given tensor field in a $2D$ grid. Or, as a secondary purpose, be used to visualize the light transport in a defined, given geometrical scene in a $2D$ grid. However, the tensor field needs to be pre-processed somehow to generate a tensor induced footprint map of transmission profiles, which we will do via matrix decompositions. For this purpose, we will let the transmission profiles act on our simple light propagation scheme to have a kind of tensor controlled directive effect on the light transport. The transmission profiles should modulate the outgoing intensities in a sense of redirection or more appropriately, redistribution. Of course, the propagation scheme will need some kind of stimulus which can be defined through symbolic functions (eg. $r(\omega)=1$) by the user which we sample as a primary step. This way, the approach will operate as a global illumination visualization technique for light distributions in polar coordinates provided the user set an initial profile into the grid. In the sense of automization and optimzation, another consequent aim is to implement an approach borrowed from vector-field visualization to reveal structures, which are hard to graps by ordinary methods. We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges representing LCS (Lagrangian Coherent Structures). These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature such as moraines in glaciers. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of varying directional distributions (manifesting in diverging/converging tensor field lines or higher tensor magnitude) represented by tensors as we use our tensor controlled light transport method as a basis or framework to compute the FTLE field. Thus, we propose another method to compute an FTLE like field on tensor fields. The FTLE map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions and since we effectively measure the finite-time Lyapunov exponent (FTLE) with both forward and backward integration time (stimulus directions, in our case), we will detect both cases. Both invented methods, will give new insights into the topology and anisotropy of tensor fields and provide tools for the structural analysis. The greater goal of this work is therefore the encoding of a tensor field into certain elaborate representations and the interpretation of meaningful results comparing the proposed techniques against others. The designed approach could e.g. be applied in medicinal sciences, engineering and physics to reveal structures which are hard to grasp through interpreting simpler and sometimes confusing glyph and tensor field line visualizations.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Conclusion and Future Work}
\label{chap:concl}
We have presented a presented a technique to extract meaningful structures from tensor fields, which can in turn also be used to visualize meaningful structures in $2D$ virtual geometrical scenes. To put it into context, we discussed its relations to vector field visualization techniques and other tensor field and light transport visualization methods. The visualization is derived from global light transport methods propagating intensities inside of a given Cartesian grid consisting of cells with given polar profiles respectively. Based on this propagation algorithm, we can compute a FTLE-like field on the tensor field.

>>>>

\chapter*{Abstract}
%\input{abstract}
Many problems in science and engineering require
tensor field representations. Tensor fields, are most commonly associated with stress distributions (cf. Cauchy-stress tensor) in 2D/3D grids, but also have some other meanings in practice in physics. By means of global illumination techniques, we motivate a new method to visualize tensor fields, which is also capable of visualizing the light transport (propagation) in a topologically defined scene geometry. As a basis, we derive a simple light propagation scheme for regular grids, which satisfies propagation attenuation and energy-conservation principles and is able to approximate light distributions for given light source position(s) and direction(s) until convergence. The transmission profiles within this grid are considered as crystal fiber structures as an underlying physical model (, eg., gemstones: tiger's eye's cat's eye effect). The consequent task is to model anisotropic fibre structures with the orientation and anisotropy measures of the underlying tensor field in the grid.
For this, we derive an eigenbasis by PCA (principal component analysis) for every cell of the tensor field and form an ellipsoid equation as transmission (transfer/weighting) function for the transmitted light profiles. Hence, the resulting light distribution can be visualized via 2D scalar field color coding or polar plots for directional information. Additionally, we measure impulse responses of the tensor field with Dirac-pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions. This map is considered as a global illumination energy flow field.
We gain inspiration by vector field particle tracing's FTLE approach for analyzing the gradient of the flow field of the resulting light distributions. Similar to how the FTLE detects ridges/LCS (Lagrangian Coherent Structures) in vector fields (flow fields), our approach captures these structures likewise, namely attracting/repelling/saddling tensor field lines (TFL) which frequently represent key structures in nature, in tensor fields. We denote this entity a light transport gradient (LTG) which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global light transport  directed by imprint/engraving in crystelline fiber structures modeled by the tensor field's ellipsoid glyhps. We evaluate our approaches for plausibility and comparability within an extensive test campaign and demonstrate its performance on several examples of datasets.
\newpage

@@

The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies, e.g., clustering, structuring or projection techniques. When it comes to tensor fields, this comes in handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor-medical resonance imaging) \cite{hlawatsch}, when measuring water molecule diffusion characteristics within tissue or in simulated (or measured) stress tensor fields from solid and fluid mechanics. In these domains of science the problem arises, that tensor fields can be generated (simulated) or measured with physical/mathematical background, but not be regarded intuitively. Necessity begets ingenuity as always, which lets visualization methods and systems emerge in this context. Concurrently, the problem setting of this work encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids consisting of mainly second-order tensors (cf. Cauchy stress tensor). These grids span a spatial domain in general with sample positions indicated by vertices. Tensor fields feature a sampling grid consistent of tensors which is spaced uniformly in the commonly used regular grid. A regular grid, which we will utilize in this context, is a normalized uniform grid as a special case of a regular grid. We aim to use global illumination methods to model light transport directed by tensor fields within such type of grids, which we consider as light propagation volumes (LPVs). Congruently, the most basic objective of this work is to design a simple light propagation scheme yet following the most basic but crucial physical principles and laws, like energy conservation and propagation attenuation, for tensor field visualization. Furthermore, it should hold the additional possibility to model light transport in a geometrical scene. This propagation scheme should either be used to visualize the light transport paths (,defined by the anisotropy of tensor fields and tensor magnitudes of a given tensor field in a $2D$ grid. Or, as a secondary purpose, be used to visualize the light transport in a defined, given geometrical scene in a $2D$ grid. However, the tensor field needs to be pre-processed somehow to generate a tensor induced footprint map of transmission profiles, which we will do via matrix decompositions. For this purpose, we will let the transmission profiles act on our simple light propagation scheme to have a kind of tensor controlled directive effect on the light transport. The transmission profiles should modulate the outgoing intensities in a sense of redirection or more appropriately, redistribution. Of course, the propagation scheme will need some kind of stimulus which can be defined through symbolic functions (eg. $r(\omega)=1$) by the user which we sample as a primary step. This way, the approach will operate as a global illumination visualization technique for light distributions in polar coordinates provided the user set an initial profile into the grid. In the sense of automization and optimzation, another consequent aim is to implement an approach borrowed from vector-field visualization to reveal structures, which are hard to graps by ordinary methods. We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges representing LCS (Lagrangian Coherent Structures). These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature such as moraines in glaciers. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of varying directional distributions (manifesting in diverging/converging tensor field lines or higher tensor magnitude) represented by tensors as we use our tensor controlled light transport method as a basis or framework to compute the FTLE field. Thus, we propose another method to compute an FTLE like field on tensor fields. The FTLE map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions and since we effectively measure the finite-time Lyapunov exponent (FTLE) with both forward and backward integration time (stimulus directions, in our case), we will detect both cases. Both invented methods, will give new insights into the topology and anisotropy of tensor fields and provide tools for the structural analysis. The greater goal of this work is therefore the encoding of a tensor field into certain elaborate representations and the interpretation of meaningful results comparing the proposed techniques against others. The designed approach could e.g. be applied in medicinal sciences, engineering and physics to reveal structures which are hard to grasp through interpreting simpler and sometimes confusing glyph and tensor field line visualizations.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Conclusion and Future Work}
\label{chap:concl}
We have presented a presented a technique to extract meaningful structures from tensor fields, which can in turn also be used to visualize meaningful structures in $2D$ virtual geometrical scenes. To put it into context, we discussed its relations to vector field visualization techniques and other tensor field and light transport visualization methods. The visualization is derived from a global light transport method propagating intensities inside of a given regular grid consisting of cells with given polar profiles respectively. Based on this propagation algorithm, we can compute a FTLE-like field on the tensor field. This field can be visualized as scalar field (e.g. heatmap representation) to extract ridges, i.e., key regions separating areas of correlating directional distributions.

A possible application would be mechanical component design, as we could segment ROIs exhibiting increased loads or in a greater scope even machine component design, which are produced in series. The invented technique could also aid geologist in analyzing stress in cross-sections of the earth to be able to predict earthquakes sufficiently soon enough.
